---
title: "Example - Session 6: Uncertainty"
author: "David Schweizer"
date: "12 10 2023"
output: 
  html_document:
    highlight: "tango"
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Coding example

## Load libraries:

```{r load-libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  "tidyverse",
  "ggridges",
  "gghalves"
)
```

## Load data:

We use data from the latest wave of the European Social Survey (Self-completion):

```{r load-data, message = FALSE, warning=FALSE}
# https://ess-search.nsd.no/en/study/172ac431-2a06-41df-9dab-c1fd8f3877e7

ESS10SC <- read_csv("C:/Users/David/Desktop/ESS10SC.csv")

# codebook: https://stessrelpubprodwe.blob.core.windows.net/data/round10/survey/ESS10SC_appendix_a7_e03_0.pdf

# idno - Person ID
# cntry - Country
# stfdem - Satisfaction with democracy
# stfgov - Satisfaction with the government
```

First, we are only interested in satisfaction with democracy in each country. We have to remove missing values in the satisfaction with democracy variable. As it ranges from 0 to 10, we can just filter all other values. The same is true for satisfaction with the government:

```{r select}
ess <- ESS10SC %>% 
  select(idno, cntry, stfdem, stfgov) %>% 
  filter(stfdem <= 10) %>% 
  filter(stfgov <= 10)
```

## Histograms

Histogram of satisfaction with democracy. Bin width of 1 and white edges of bars.

```{r histogram1}
ggplot(ess, 
       aes(x = stfdem)) +
  geom_histogram(binwidth = 1, 
                 color = "red")
```

Forcing the bins to start at whole numbers. Therefore, we use the `boundary` argument.

```{r histogram2}
ggplot(ess, aes(x = stfdem)) +
  geom_histogram(binwidth = 1, 
                 color = "white", 
                 boundary = 5) 
```

Now, we want to have a look at the distribution of stfdem by country:

```{r distribution-country}
ggplot(ess, aes(x = stfdem, 
                fill = cntry)) +
  geom_histogram(binwidth = 1, 
                 color = "white", 
                 boundary = 1) 
```

This graph is very colorful, but it’s basically impossible to interpret. We should not only use  filling, but also facets by month. This way, we generate separate graphs for each country. Therefore, we can turn off the fill legend as it has become redundant.

```{r distribution-country-facet}
ggplot(ess, aes(x = stfdem, 
                fill = cntry)) +
  geom_histogram(binwidth = 1, 
                 color = "white", 
                 boundary = 1) +
  guides(fill = "none") +
  facet_wrap(vars(cntry))
```

## Density plots

To create a density plot, our code is nearly identical to what we used for the histogram. We only change the `geom` layer:

```{r density-plot}
ggplot(ess, aes(x = stfdem)) +
  geom_density(color = "grey20", fill = "grey80")
```

We can change some of the calculus options if we want:

```{r density-plot-options1}
ggplot(ess, aes(x = stfdem)) +
  geom_density(color = "grey20", 
               fill = "grey80",
               bw = 0.5, 
               kernel = "rectangular")
```

Another way might be to fill by country. It is necessary to make the different layers transparent to a certain degree to see through the whole stack.

```{r density-plot-options2}
ggplot(ess, 
       aes(x = stfdem, 
           fill = cntry)) +
  geom_density(alpha = 0.5)
```

Similarly to our histogram, this is difficult to interpret. Therefore, we use facets again:

```{r density-plot-options3}
ggplot(ess, 
       aes(x = stfdem, fill = cntry)) +
  geom_density(alpha = 0.5) +
  guides(fill = "none") +
  facet_wrap(vars(cntry))
```

Furthermore, we can stack the (density) plots behind each other using `ggridges`. For that to work, we need to map `cntry` to the y-axis.

```{r density-plot-options4}
ggplot(ess, aes(x = stfdem, 
                             y = cntry, 
                             fill = cntry)) +
  geom_density_ridges() +
  guides(fill = "none")
```


Now, we can add further information to `geom_density_ridges()` with some other arguments like `quantile_lines`. We can use the quantiles argument to tell the plow how many parts to be cut into. Since we just want to show the median, we’ll set that to 2 so each density plot is divided in half:

```{r density-plot-options5, message=FALSE}
ggplot(ess, aes(x = stfdem, 
                             y = cntry, 
                             fill = cntry)) +
  geom_density_ridges(quantile_lines = TRUE, 
                      quantiles = 2) +
  guides(fill = "none")
```

Now that we have good working code, it is easy to substitute in other variables by changing the x mapping:

```{r density-plot-options6, message=FALSE}
ggplot(ess, aes(x = stfgov, 
                y = cntry, 
                fill = cntry)) +
  geom_density_ridges(quantile_lines = TRUE, 
                      quantiles = 2) +
  guides(fill = "none")
```


## Box, violin, and rain cloud plots

In addition, we can look at the distribution of variables with box plots, violin plots, etc. 

First, we’ll make a box plot of satisfaction with democracy:

```{r boxplot}
ggplot(ess,
       aes(y = stfdem)) +
  geom_boxplot()
```

We can switch this to a violin plot by just changing the geom layer:

```{r violinplot}
ggplot(ess,
       aes(y = stfdem,
           x = "")) +
  geom_violin()
```

With violin plots it’s typically good to overlay other geoms. We can add some jittered points for a strip plot:

```{r violinplot-overlay1, warning = FALSE}
ggplot(ess,
       aes(y = stfdem,  
           x = "")) +
  geom_violin() +
  geom_point(size = 0.2, position = position_jitter(width = 0.1)) +
  guides(fill = "none")
```

Overlaying the points directly on top of the violins shows extra information. However, it is also more crowded and more difficult to read. We can use the `gghalves` package to generate special halved versions of some of these geoms:

```{r gghalves-box, warning = FALSE}
ggplot(ess,
       aes(y = stfdem,  
           x = "")) +
  geom_half_point(side = "l", size = 0.5) +
  geom_half_boxplot(side = "r")
```

Note the side argument for specifying which half of the column the geom goes. We can also use geom_half_violin():

```{r gghalves-violin, warning = FALSE}
ggplot(ess,
       aes(y = stfdem,  
           x = "")) +
  geom_half_point(side = "l", size = 0.5) +
  geom_half_violin(side = "r")
```


And finally, if we flip the plot, we can  generate a rain cloud plot:

```{r rain-cloud-plot, warning =FALSE}
ggplot(ess,
       aes(y = stfdem,  
           x = "")) +
  geom_half_boxplot(aes(alpha = 0.2), side = "l", width = 0.05, nudge = -0.15) +
  geom_half_point(side = "l", size = 0.2) +
  geom_half_violin(side = "r", nudge = 0.2) +
  coord_flip()
```
```


# Your turn!

Adapt the code to 

- look at satisfaction with democracy in Germany for different subgroups you are interested in. This could include, for example, gender, party identification, political ideology, ...

- create a raincloud plot for only Germany.

- get the ESS data from earlier waves and compare satisfaction with democracy (or any other variable) with ridgeplots over time!

- or use the code to visualize uncertainty in a data set of your choice!




