---
title: "session_08_merge_transfrom"
author: "David Schweizer"
date: "2023-10-26"
output: 
  html_document:
    highlight: "tango"
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This script

This script demonstrates how to merge data from different sources and transform it. We first load the data from [The Populist](https://popu-list.org/) and [ParlGov](http://www.parlgov.org/#data). Second, we then transform the data and merge it.

The goal is to plot the vote share *- we get this from the ParlGov data -* of Eurosceptic parties *- we get this from The Populist -* from 1989 until 2022 for West European countries.

In the section **Your turn: Add manifesto data**, you can apply the learnings and merge our dataset with data from the *Manifesto Project*.

# Setup

The following code chunk makes sure that we load all the necessary packages. If they are not installed, the function installs and loads the packages for us.

```{r load-packages}
# load your packages here
pacman::p_load("tidyverse")
```

# Load data

We first load the Populist data. We can load the data from our device, however, we can also download the data from the online source:

```{r load-populist-data}
# we use the read.csv2 function because the populist data is separated with a ";"
populist <- read.csv2("https://osf.io/download/7krzm/")
```

Next, we load the parlgov data:

```{r load-parlgov-data}
# here, we use the read.csv function because the separator is a ","
parlgov <- read.csv("https://www.parlgov.org/data/parlgov-development_csv-utf-8/view_election.csv")
```

# Inspect data

We have loaded the data. Now, we should check the data structure and variables:

```{r check-populist}
str(populist)
```

```{r check-parlgov}
str(parlgov)
```

You can see that the populist dataframe has a column called parlgov_id and our parlgov dataframe has a column called party_id. This is how we can merge the two data frames later!

However, in this script, we are only interested in West European countries.

# Transform

-   First, we select only West European countries.

-   Second, the Populist starts in 1989. Therefore, we only want to keep elections in or after 1989.

-   Third, we are only interested in national elections. We discard European elections.

```{r select-west-european-countries}
# we create a new dataframe called parlgov_west
parlgov_west <- parlgov %>% 
  # we filter the parlgov dataframe to only keep these West European countries
  filter(country_name %in% c("Germany",
                        "Austria",
                        "Switzerland",
                        "Belgium",
                        "Netherlands",
                        "France",
                        "Spain",
                        "Portugal",
                        "United Kingdom",
                        "Ireland")) %>% 
  # we only keep observations in or after 1989
  filter(election_date >= "1989-01-01") %>% 
  # we only keep national elections
  filter(election_type == "parliament")
  
```

-   Now, we merge the two datasets. We use the *left_join* function*. <https://dplyr.tidyverse.org/reference/mutate-joins.html>*

```{r merge}
# we create a new data frame called parl_pop
parl_pop <- parlgov_west %>% 
  # we join with the populist dataframe:
  # we keep all observations in parlgov west, and add all rows with matching party ids from the populist dataframe!
  left_join(populist, by = c("party_id" = "parlgov_id"))

# have a look at our new dataframe
# there are many NA's because many parties are not listed in the Populist as they are not Eurosceptic, far-right, populist, etc.
head(parl_pop)
```

-   The populist provides a dynamic classification of parties. Therefore, we have to check, if a party is actually Eurosceptic at the particular election.

```{r account-for-dynamic-classification}
# we "overwrite" the parl_pop dataframe
parl_pop <- parl_pop %>% 
  # we create a variable called "Eurosceptic" that takes the value "1" if 
  # it the party was considered eurosceptic before and after the election
  mutate(Eurosceptic = case_when(election_date >= eurosceptic_start & election_date <= eurosceptic_end ~ 1,
                                 TRUE ~ 0))
```

-   Now, we can restrict the dataset to the variables we are interested in for plotting!

```{r select-relevant-variables}
# we create a new dataframe
parl_pop_plot <- parl_pop %>% 
  # we select only the variables we need for our first plot
  # and rename some of them in one step!
  # dplyr:select(new_name = old_name)
  dplyr::select(country = country_name.x,
                election_date,
                vote_share,
                Eurosceptic) 
  
```

-   As a last step, we need to summarize the vote shares of all Eurosceptic parties for each election.

```{r summarize-vote-shares}
# We "overwrite" our existing data frame
parl_pop_plot <- parl_pop_plot %>% 
  # before we summarize, we need to group our variables
  group_by(country, election_date, Eurosceptic) %>% 
  # now we can summarize, na.rm = TRUE omits NAs during the calculation
  # this is important as some vote_shares are missing
  summarize(Vote_share = sum(vote_share, na.rm = TRUE))
```

# Plot

Finally, we can plot the vote share of Eurosceptic parties in West European countries over time!

```{r basic-plot}
# here, we are only interested in the vote share of Eurosceptic parties
# We create a new dataframe called "Eurosceptic_vote"
Eurosceptic_vote <- parl_pop_plot %>% 
  # and filter that we keep the Eurosceptic observations
  filter(Eurosceptic == 1) %>% 
  # and change election_date to date format
  mutate(election_date = as.Date(election_date))

# Now we can plot these vote shares of Eurosceptic parties
# we need out Eurosceptic_vote dataframe
ggplot(data = Eurosceptic_vote,
       # election_date at the x axis to show the development over time
       aes(x = election_date,
           # vote share on the y axis
           y = Vote_share,
           # we color by country
           color = country,
           # and tell ggplot that we need to group the observations by country
           group = country)) +
  geom_line() +
  theme_minimal()
```

Is a smoothed trend better?

```{r smoothed}
ggplot(data = Eurosceptic_vote,
       aes(x = election_date,
           y = Vote_share,
           color = country,
           group = country)) +
  # we set se=F to avoid confidence intervals as they would overlap too much
  geom_smooth(se = FALSE) +
  theme_minimal()
```

Small multiples?

```{r small-multiples}
# how would you produce this plot?
ggplot(data = Eurosceptic_vote,
       aes(x = election_date,
           y = Vote_share)) +
  geom_line() +
  # we can just use the facet_wrap function to produce these small multiples!
  facet_wrap(vars(country)) +
  theme_minimal()

```

How could we make small multiples with sparklines?

```{r small-multiples-with-sparklines}
# how would you produce this plot?
ggplot(data = Eurosceptic_vote,
       aes(x = election_date,
           y = Vote_share)) +
  geom_line() +
  # again, we use facet_wrap, scales allows us that the y-axis can vary, ncol=3 just defines that we want to facet in 3 columns
  facet_wrap(vars(country), scales = "free_y", ncol= 3) +
  # theme_void gives us this "sparkline" look. We now can only concentrate on the trend. No labels etc.
  theme_void() +
  # this makes the header for each facet in bold text
  theme(strip.text = element_text(face = "bold"))
```

# Your turn: Add manifesto data

-   First, you have to load the manifesto data. You can get the .csv file from ILIAS. Alternatively, you can create a free account here:<https://manifesto-project.wzb.eu/datasets/MPDS2023a>

-   In addition, you need to load the party information from the ParlGov database. This is the right file: <https://www.parlgov.org/data/parlgov-development_csv-utf-8/view_party.csv>

```{r load-new-data}
# load the parlgov party and the manifesto data here
parlgov_parties <- read.csv("https://www.parlgov.org/data/parlgov-development_csv-utf-8/view_party.csv")
CMP <- read.csv("MPDataset_MPDS2023a.csv") 
```

-   Next, you have to get the identifier from the Parlgov party data (cmp), then merge it to the the existing parlgov election data. This allows you to merge the new dataframe with the Manifesto data!

    -   Think first about what identifiers you need to merge theses different dataframes!

```{r get-cmp-code-from-parlgov-and-merge}
# first, we need the identifier in the parlgov data to merge our exisiting dataframe with the manifesto data! This column is called "cmp". You can only know this by inspecting the codebook!

# we create a new dataframe called parlgov_parties
parlgov_parties <- parlgov_parties %>% 
  # here, we are only interested in "party_id" to merge with our existing dataframe, and in "cmp" which allows us to link the existing dataframe including the parlgov election data and the populist data with the data from the Manifesto project.
  select(party_id, cmp)

# now we create the new dataframe called "parl_pop_manifesto": 
# we start with the existing dataframe parl_pop (this includes the parlgov election data and the populist data)
parl_pop_parties <- parl_pop %>% 
  # we then join the parlgov_parties dataframe which includes only the parlgov party identifier and the identifier for the Manifesto data (cmp)
  # we can left_join by only party_id because it is both in the parl_pop and the parlgov parties dataframe
  left_join(parlgov_parties, by = c("party_id"))  %>%
  # now, we could theoretically join the Manifesto data. However, we need another identifier! 
  # in both dataframes, the party identifier appears multiple times. Therefore, it is not clear how to match!
  # we need a second identifier: the election date - we observe at each election the party once. 
  # this way, we have a unique identification. However, we need to make sure, that the election data has the same format in both dataframes.
  # it is easiest, to transform both election date variables to the date format:
  mutate(election_date = as.Date(election_date))

# here, we transform the election date (edate) in the Manifesto Project data
CMP <- CMP %>% 
  # Note, that the date is provided in a different format here!
  # We have to tell the as.Date() function the structure: "%d/%m/%Y" -> this means for example: dd.mm.YYYY 
  mutate(edate = as.Date(edate,  format = "%d/%m/%Y"))

# finally we can join the data frames
parl_pop_CMP <- parl_pop_parties %>% 
  # the party identifier in the CMP dataframe is called "party"
  left_join(CMP, by = c("cmp" = "party", "election_date" = "edate"))
```

-   Now, you can proceed as before. Select the relevant variables and plot the salience and positions regarding European integration for the Eurosceptic parties. Here, you can see how these variables are defined by the Manifesto project: <https://manifesto-project.wzb.eu/information/documents/visualizations>

    -   We can add up the variables per108 and per110 to get a salience measure of European Integration (Salience): `per108 + per110`

    -   We can substract per110 from per108 to get a positional measure of European Integration (Position): `per108 - per110`

```{r select-and-transform-relevant-data}
# select and transform relevant data
# we "overwrite" the parl_pop_CMP dataframe
parl_pop_CMP <- parl_pop_CMP %>% 
  # and only selet the following variables
  select(country = country_name.x, election_date, Eurosceptic, vote_share, per108, per110) %>% 
  # then, we create the positional and salience measure of European integration:
  # we get the positional measure by substracting: per108 - per110
  mutate(position_EU = per108 - per110,
          # we get the salience measure by adding: per108 + per110
         salience_EU = per108 + per110) %>% 
  # now, we summarize the measures for the Eurosceptic parites, this is the same procedure as for the vote share before!
    # before we summarize, we need to group our variables
  group_by(country, election_date, Eurosceptic) %>% 
  # now we can summarize, na.rm = TRUE omits NAs during the calculation
  # this is important as some positions might be missing
  summarize(position_EU = sum(position_EU, na.rm = TRUE),
            salience_EU = sum(salience_EU, na.rm = TRUE))
  
```

-   Finally, we can plot for example the positions regarding European integration from Eurosceptic parties over time.

-   First, the position:

    -   **! Do you see any strange patterns?**

```{r plot}
# again, we are only interested in Eurosceptic positions. Therefore, we filter accordingly:
parl_pop_CMP_Eurosceptic <- parl_pop_CMP %>% 
  filter(Eurosceptic == 1)

# instead of vote share, we want to show the position regarding European integration on the y axis!
ggplot(data = parl_pop_CMP_Eurosceptic,
       aes(x = election_date,
           y = position_EU)) +
  geom_line() +
  # we can just use the facet_wrap function to produce these small multiples!
  facet_wrap(vars(country)) +
  theme_minimal()
```

-   **As you can see, the aggregated position for France are always zero. The same is true for Germany and Belgium after some time. What went wrong?**

    -   Unfortunately, the party identifiers in the parlgov data set seem not to be updated regularly to match the Manifesto party identifiers. You can read more up on this here: <https://manifesto-project.wzb.eu/tutorials/parlgov_merge>

    -   This means that we would have to adapt the party identifiers of Eurosceptic parties in our parl_pop_parties dataframe before we merge it with the CMP dataframe.

    -   For France, we have a different problem. There are two rounds of elections. Unfortunately, the parlgov and Manifesto project use different rounds as election dates. One easy solution might be to transform our data variable to just the year format (e.g., YYYY). This way, we can avoid the problem that the exact dates do not match.

-   **This is a great example that it is essential that we always check our data!**

-   As an example, let's correct the German data together. How would I tackle such a case?

    -   First: Google! This would lead me to the already mentioned blogpost: <https://manifesto-project.wzb.eu/tutorials/parlgov_merge>. Here, I could get the right code.

    -   What if Google does not help? We need to check it in detail:

```{r check-detail}
# first, we want to have a look at the Eurosceptic German parties before we merge with the CMP dataframe
# this is: parl_pop_parties

parl_pop_parties %>% 
  filter(country_name.x == "Germany",
         Eurosceptic == 1) %>% 
  select(country_name.x, election_date, party_name.x, cmp)

```

-   We can see that "Die Republikaner" and the "AfD" have no cmp identifier! Also we know that the cmp identifier for The Left is 41221. Now, we need to check the party id in the Manifesto data frame:

```{r check-Manifesto-dataframe}
CMP_test <- CMP %>% 
  filter(countryname == "Germany",
         edate >= 1990-12-02) %>% 
  # the party identifier is here called "party"
  select(countryname, edate, partyname, party) 

# check the unique party names in the CMP_test dataframe
unique(CMP_test$partyname)
```

-   We can see that the Manifesto data includes these parties: "The Left. Party of Democratic Socialism", "The Left", and the "Alternative for Germany". However, it does not include "Die Republikaner". Let's check now the party identifier for the Left party:

```{r check-party-ids}
CMP_test %>% 
  filter(partyname %in% c("Alternative for Germany", 
                          "The Left",
                          "The Left. Party of Democratic Socialism")) %>% 
  select(partyname, party) %>% 
  unique()
```

-   We see that the identifiers for the Left and its predecessor are 41222 and 41223. In parlgov it is 41221!

-   The identifier for the Alternative for Germany is 41953.

-   One approach to fix our problem with the German data can be to change the party identifier for the AfD in our "parl_pop_parties" dataframe accordingly. And the party identifier in the Manifesto dataframe for the variations of the Left. We can do this easily with the case_when() function:

```{r update-party-ids}
# we "overwrite" our parl_pop_parties dataframe
parl_pop_parties <- parl_pop_parties %>% 
  # and update the cmp column:
  # when party_name.x equals "Alternative für Deutschland" we replace the value with 41953
  # otherwise (TRUE) we keep the original cmp value.
  mutate(cmp = case_when(party_name.x == "Alternative für Deutschland" ~ 41953,
                         TRUE ~ cmp))

# now we change the identifiers for the Left party variations in the Manifesto dataframe:
CMP <- CMP %>% 
  # when party is either "The Left " or "The Left. Party of Democratic Socialism" then replace with 41221
  # Otherwise: keep the original party identifier!
  mutate(party = case_when(partyname == "The Left" | partyname == "The Left. Party of Democratic Socialism" ~ 41221,
                           TRUE ~ party))
```

-   After updating the identifiers, we should be able to run the code from above, merge the data and plot it!

```{r check-again}
parl_pop_CMP <- parl_pop_parties %>% 
  # the party identifier in the CMP dataframe is called "party"
  left_join(CMP, by = c("cmp" = "party", "election_date" = "edate"))

# select and transform relevant data
# we "overwrite" the parl_pop_CMP dataframe
parl_pop_CMP <- parl_pop_CMP %>% 
  # and only selet the following variables
  select(country = country_name.x, election_date, Eurosceptic, vote_share, per108, per110) %>% 
  # then, we create the positional and salience measure of European integration:
  # we get the positional measure by substracting: per108 - per110
  mutate(position_EU = per108 - per110,
          # we get the salience measure by adding: per108 + per110
         salience_EU = per108 + per110) %>% 
  # now, we summarize the measures for the Eurosceptic parites, this is the same procedure as for the vote share before!
    # before we summarize, we need to group our variables
  group_by(country, election_date, Eurosceptic) %>% 
  # now we can summarize, na.rm = TRUE omits NAs during the calculation
  # this is important as some positions might be missing
  summarize(position_EU = sum(position_EU, na.rm = TRUE),
            salience_EU = sum(salience_EU, na.rm = TRUE))

# again, we are only interested in Eurosceptic positions. Therefore, we filter accordingly:
parl_pop_CMP_Eurosceptic <- parl_pop_CMP %>% 
  filter(Eurosceptic == 1)

# instead of vote share, we want to show the position regarding European integration on the y axis!
ggplot(data = parl_pop_CMP_Eurosceptic,
       aes(x = election_date,
           y = position_EU)) +
  geom_line() +
  # we can just use the facet_wrap function to produce these small multiples!
  facet_wrap(vars(country)) +
  theme_minimal()
```

-   **Nice! We managed to plot the "correct" data for Germany.** However, we already know one limitation. The Republicans are not included in the Manifesto Project. Though, one has to note, that the party is not significant in recent years from an electoral perspective.

# Add on

-   You can try to correct the data for France and Belgium if you want to.

```{=html}
<!-- -->
```
-   What about the salience and positions regarding European integration from non-Eurosceptic parties?

-   Is there a correlation between vote shares and salience/positions regarding European integration?
