---
title: "Public Opinion on Climate Change"
subtitle: "Week4"
author: "David Schweizer"
date: "March 07, 2023"
date-format: long
institute: "University of Mannheim"
format: 
  revealjs:
    theme: serif
    slide-number: c/t #< collapsed/total
    footer: "The Politics of Climate Change | Spring 2023 | University of Mannheim"
editor: visual
---

## Plan for Today

-   Concepts so far & Discussion of readings \~ 25 min

-   First research question \~ 45 min

    -   Research question / puzzle?

    -   Literature?

    -   Data?

-   Discussion of research questions \~ 20 min

-   Outlook next week

## Mechanisms, Dilemmata, Politics

::: panel-tabset
#### Pathways

**Direct**

-   Change in distribution of weather leads to change in distribution of events experienced by individuals

**Informational**

-   Individuals expect specific distributions of weather events and their effects

#### Dilemmata

**Collective action theory:**

-   Incentives for free-riding need to be overcome by international cooperation

**Distributive conflict:**

-   Climate policies â¤ New economic winners and losers
-   Conflicts over material benefits, reinforced by (ideological) political struggles

#### Energy transition

*Experience* or *learning curves:* Often, policy and politics not taken into account

-   Affects both price and quantity
-   Effects differ depending on current stage of the experience curve
:::

## Quiz

<iframe src="https://app.sli.do/event/4X6Tgk5jB9pr61Ny4a46br" height="80%" width="100%" frameBorder="0" style="min-height: 450px;" title="Slido">

</iframe>

## Quiz

Respondents in Germany were asked: \[...\] *what percentage of climate scientists have concluded that human-caused climate change is happening?* What is their average estimate?

-   ::: {.fragment .highlight-red}
    69%
    :::

-   79%

-   89%

-   99%

## Quiz

Respondents in Germany were asked whether they think that *oil companies are hiding technology that could make cars run without petrol or diesel*. How many think this is the case?

-   23%

-   ::: {.fragment .highlight-red}
    46%
    :::

-   64%

-   80%

## Quiz

Respondents in Germany were asked when, if ever, they think *climate change will start to harm them personally*. How many think it already harms them or in the next ten years?

-   33%

-   ::: {.fragment .highlight-red}
    43%
    :::

-   53%:

-   63%

## Quiz

And how many respondents strongly or tend to agree with the statement that *there is no point in changing their behaviour to tackle climate change because it won't make any difference anyway*.

-   15%

-   ::: {.fragment .highlight-red}
    25%
    :::

-   35%

-   45%

## Surveys

-   PERITIA (2022). Public perceptions on climate change. [Link](https://peritia-trust.eu/wp-content/uploads/2022/06/4-Climate-change_EU.pdf)
-   Alliance of Democracies (2022). Democracy Perception Index 2022. [Link](https://docs.google.com/spreadsheets/d/1gdjTHqM_O2kDNyFgvLWVsCn4lFWzFRoMRUdTelXxTM4/edit#gid=1271546525)
-   European Investment Bank (2022). 2021-2022 EIB Climate Survey. [Link](https://www.eib.org/en/surveys/climate-survey/4th-climate-survey/index.htm)
-   Yale Program on Climate Change Communication and Data for Good at Meta (2022). [Link](https://climatecommunication.yale.edu/publications/international-public-opinion-on-climate-change-2022/)

## Bush & Clayton

. . .

-   Gender gap in wealthy but not in poor countries

    -   Men tend to attach greater material costs to mitigation

    -   Men tend to express more support for group-based dominance (correlated with climate skepticism)

-   Implications

    -   Women's greater role in climate policy may lead to more decisive climate action

    -   Voices and perspectives from the Global South need to be considered

## Lee et al.

. . .

-   Predictors of public climate change awareness:

    -   Educational attainment

-   Predictors of climate risk perception:

    -   Understanding the anthropogenic cause of climate change (Europe, Latin America)

    -   Perception of local temperature change (Africa, Asia)

Improving education, climate literacy, and understanding of the local impacts are essential for more support for climate action - *exact strategy depends on country (context)*.

## Discussion Questions

-   Would female politicians (head of states) implement more progressive climate change mitigation policy?

-   Educational attainment as the strongest predictor of climate awareness. Climate skeptics just uneducated?

-   How could climate education look like? What should we teach over one school year, for example? Climate education as a political issue?

-   What role do social media / modern communication and globalization play? How do they foster the influence of right-wing populist / climate change denying parties?

## Finding a Research Question

**Task**

In small groups or individually, find a research question that you could also follow-up with your term paper.

-   Academic relevance?

-   Societal relevance?

-   Feasibility?

-   Relation to existing theories?

Formulate the question with a question mark!

## Finding Literature

**Helpful tools**

-   Elicit [Link](https://elicit.org/)
-   Research Rabbit [Link](https://researchrabbitapp.com/home)
-   Connected Papers [Link](https://www.connectedpapers.com/)

## Finding Data

**Task**

In small groups or individually, find possible data sources for answering your research question.

-   Easily accessible?

-   What kind of data? E.g., cross-sectional, panel data?

-   What concepts can you measure with these data?

## Your Turn!

If you have absolutely no idea where to start:

-   Have a look at the surveys, anything interesting there?

-   Recent, climate change-related events? Reporting in newspapers, etc.

```{r countdown-package, include=FALSE}
#remotes::install_github("gadenbuie/countdown")
library(countdown)
```

```{r, echo = FALSE}
countdown(minutes = 45, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

# Discussion of Research Questions

## (Until) Next Week

-   BLOCK II - Political Behaviour (2): Natural Disasters, Climate Attitudes and Voting Behaviour. Add comments, discussion questions, etc. for the mandatory readings. Try Perusall:

    -   Garside, S., & Zhai, H. (2022). If not now, when? Climate disaster and the Green vote following the 2021 Germany floods. Research & Politics, 9(4).

    -   Hilbig, H., & Riaz, S. (2021). Natural Disasters and Electoral Support for Green Parties-Evidence from the 2021 Flood in Germany. OSF Preprints.
