---
title: "Populist Backlash"
subtitle: "Week10"
author: "David Schweizer"
date: "May 02, 2023"
date-format: long
institute: "University of Mannheim"
format: 
  revealjs:
    theme: serif
    slide-number: c/t #< collapsed/total
    footer: "The Politics of Climate Change | Spring 2023 | University of Mannheim"
editor: visual
---

## Plan for Today

-   Presentations & discussion questions

-   Discussion format:

    -   When discussing the question, also keep in mind the readings that have been covered so far.

    1.  Individual thinking: 2 min.

    2.  Discussing in small groups (of 2-3 persons): 4 min.

    3.  Sharing your thoughts in the plenum: \~10 min.

## Presentation 1

-   McCright, A. M., & Dunlap, R. E. (2011). The politicization of climate change and polarization in the American public's views of global warming, 2001-- 2010. The Sociological Quarterly, 52(2), 155-194

## Discussion - Individual Thinking

-   McCright & Dunlap find a political divide between democrats and republicans in the American public. But what about multi-party systems? Divide based on left-right or cutting across the party spectrum?

```{r countdown-package, include=FALSE}
#remotes::install_github("gadenbuie/countdown")
library(countdown)
```

```{r, echo = FALSE}
countdown(minutes = 2, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - In Groups

-   McCright & Dunlap find a political divide between democrats and republicans in the American public. But what about multi-party systems? Divide based on left-right or cutting across the party spectrum?

```{r, echo = FALSE}
countdown(minutes = 4, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - Sharing

<iframe src="https://app.sli.do/event/7ZcX6uRuWLPenmFjRTWvV6" height="100%" width="100%" frameBorder="0" style="min-height: 560px;" title="Slido">

</iframe>

## Presentation 2

-   Forchtner, B. (2019). Climate change and the far right. Wiley Interdisciplinary Reviews: Climate Change, 10(5), e604.

## Discussion - Individual Thinking

-   How do right-wing parties that deny climate change react to "climate refugees" and how does society act?

```{r, echo = FALSE}
countdown(minutes = 2, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - In Groups

-   How do right-wing parties that deny climate change react to "climate refugees" and how does society act?

```{r, echo = FALSE}
countdown(minutes = 4, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - Sharing

<iframe src="https://app.sli.do/event/7ZcX6uRuWLPenmFjRTWvV6" height="100%" width="100%" frameBorder="0" style="min-height: 560px;" title="Slido">

</iframe>

## Presentation 3

-   Huber, R. A., Fesenfeld, L., & Bernauer, T. (2020). Political populism, responsiveness, and public support for climate mitigation. Climate Policy, 20(3), 373-386.

## Discussion - Individual Thinking

-   Huber et al. show that populist attitudes enhance the effect of partisanship on climate concern/support for climate policies in the US. What are the implications to address climate change in other countries, particularly with high levels of populism?

```{r, echo = FALSE}
countdown(minutes = 2, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - In Groups

-   Huber et al. show that populist attitudes enhance the effect of partisanship on climate concern/support for climate policies in the US. What are the implications to address climate change in other countries, particularly with high levels of populism?

```{r, echo = FALSE}
countdown(minutes = 4, seconds = 00, 
          left = 0, right = 0,
          padding = "25px",
          margin = "5%",
          font_size = "5em",
          #color_border = solarized$magenta,
          color_text = "#000",
          color_running_background  = "#f0f1eb",
          #color_running_text        = solarized$base02,
          color_finished_background = "#f0f1eb")
          #color_finished_text       = solarized$base3)
```

## Discussion - Sharing

<iframe src="https://app.sli.do/event/7ZcX6uRuWLPenmFjRTWvV6" height="100%" width="100%" frameBorder="0" style="min-height: 560px;" title="Slido">

</iframe>
